---
title: "Example 1: Regional inequality analysis"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```



```{r}
#| label: setup
#| context: setup
library(cvdprevent)

# We will also be usng the following packages in this article.
# These will be fully namespaced, e.g. dplyr::select() throughout.

# library(dplyr)
# library(stringr)
# library(reactable)
# library(tidyr)
# library(gt)
# library(plotly)
```

This article walks through a complete, reproducible workflow for analysing regional variation in CVD indicators using the {cvdprevent} package.

+----------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------+
| 🎯 Goal              | Identify variation in CVD indicators (e.g., AF prevalence, hypertension treatment) across a geography, such as ICS, PCN or practice to target interventions. |
+----------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------+
| 📋 Steps             | 1.  Identify time period and geography.                                                                                                                      |
|                      | 2.  Explore available indicators.                                                                                                                            |
|                      | 3.  Select an appropriate indicator.                                                                                                                         |
|                      | 4.  Retrieve and interpret results.                                                                                                                          |
+----------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------+
| 🪧 Why {cvdprevent}? | Tidyverse-friendly outputs and helper functions for indicators, time periods and areas simplify pulling and comparing NHS CVDPREVENT audit data.             |
+----------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------+

## Step 1: Select our system level and time period

Data are published quarterly, each with a unique numeric ID. The first step is to identify the relevant time period and system level. Here we filter for NHS Region-level data.

```{r}
#| label: regional inequality - time period and system level
cvd_time_period_system_levels() |>
  # keep time periods where regional data is available
  dplyr::filter(SystemLevelName == "Region") |>
  # order the time periods most recent first
  dplyr::arrange(dplyr::desc(TimePeriodID)) |>
  # select variables that appear interesting
  dplyr::select(dplyr::any_of(c(
    "TimePeriodID",
    "TimePeriodName",
    "SystemLevelID",
    "SystemLevelName"
  ))) |>
  # display in an interactive table
  reactable::reactable()
```

From this table, *TimePeriodID* value `26` ("To June 2025") is appropriate. NHS Region corresponds to *SystemLevelID* `6`.

## Step 2: Identifying available indicators

We next explore which indicators are reported for this time period and system level:

```{r}
#| label: regional inequality - indicator selection

inds <-
  cvd_indicator_list(time_period_id = 26, system_level_id = 6) |>
  # select relevant columns to help
  dplyr::select(dplyr::any_of(c(
    "IndicatorID",
    "IndicatorCode",
    "IndicatorShortName",
    "IndicatorOrder"
  )))

inds |>
  # display in an interactive table
  reactable::reactable(searchable = TRUE, defaultPageSize = 5)
```

This returns `r nrow(inds)` indicators - too many to review manually. To focus, we filter for hypertension-related measures:

```{r}
#| label: regional inequality - filter for hypertension indicators

# show hypertension-related indicators
inds |>
  dplyr::filter(stringr::str_starts(
    string = IndicatorShortName,
    pattern = "Hypertension"
  )) |>
  # sort them by the IndicatorOrder
  dplyr::arrange(IndicatorOrder) |>
  # display in a table and remove default row shading
  gt::gt() |> gt::tab_options(quarto.disable_processing = TRUE)
```

## Step 3: Obtain indicator metadata

Hypertension indicators include prevalence, monitoring, treatment to threshold and potential over-treatment. Suppose we are interested in treatment to appropriate threshold (*IndicatorID* = `32`). To understand its definition and calculation:

```{r}
#| label: regional inequality - indicator details

# gather some details about this indicator
ind_details <- cvd_indicator_details(indicator_id = 32)

# nest the details to break up into useful chunks
ind_details <-
  ind_details |>
  dplyr::select(dplyr::any_of(c(
    "MetaDataCategoryID",
    "MetaDataTitle",
    "MetaData"
  ))) |>
  tidyr::nest(data = dplyr::any_of(c("MetaDataTitle", "MetaData")))
```

::: panel-tabset
## Meta data 1

```{r}
#| label: regional inequality - indicator details - p1
ind_details[[2]][[1]] |>
  gt::gt() |>
  gt::tab_options(quarto.disable_processing = TRUE)
```

## Meta data 2

```{r}
#| label: regional inequality - indicator details - p2
ind_details[[2]][[2]] |>
  gt::gt() |>
  gt::tab_options(quarto.disable_processing = TRUE)
```

## Meta data 3

```{r}
#| label: regional inequality - indicator details - p3
ind_details[[2]][[3]] |>
  gt::gt() |>
  gt::tab_options(quarto.disable_processing = TRUE)
```

## Meta data 4

```{r}
#| label: regional inequality - indicator details - p4
ind_details[[2]][[4]] |>
  gt::gt() |>
  gt::tab_options(quarto.disable_processing = TRUE)
```
:::

From the *Definition* section, this indicator measures the percentage of people with hypertension whose blood pressure is controlled to recommended thresholds, varying by age and setting. Metadata also explains:

-   Purpose and interpretation

-   Numerator and denominator definitions

-   Suppression rules

-   Guidance for use in dashboards

## Step 4: Gather performance measures

We now have the key components:

-   **System level:** NHS Region (`system_level_id = 6`)

-   **Time period:** To June 2025 (`time_period_id = 26`)

-   **Indicator:** Hypertension treatment to threshold (`indicator_id = 32`)

-   **Metadata:** Definitions and interpretation guidance

With these, we can retrieve the raw data:

```{r}
#| label: regional inequality - indicator data - gather
ind_data <- 
  cvd_indicator_raw_data(time_period_id = 26, system_level_id = 6, indicator_id = 32)
```

This dataset provides the values needed to compare NHS Regions and assess inequality in hypertension management.

```{r}
#| label: regional inequality - indicator data - visualise
#| code-fold: true

ind_data |> 
  # get the overall performance metric
  dplyr::filter(MetricCategoryName == "Persons") |> 
  # wrangle data
  dplyr::mutate(
    # sort areas by their value
    AreaName = AreaName |> forcats::fct_infreq(w = Value),
    # prepare some labels
    lbl_value = scales::percent(Value, accuracy = 0.1, scale = 1),
    lbl_lcl = scales::percent(LowerConfidenceLimit, accuracy = 0.1, scale = 1),
    lbl_ucl = scales::percent(UpperConfidenceLimit, accuracy = 0.1, scale = 1),
    lbl_num = scales::comma(Numerator),
    lbl_denom = scales::comma(Denominator),
    # create text to show the overall value on the bar
    Text = lbl_value |> as.character(),
    # create some text to include in the context box when you hover over the plot
    HoverText = glue::glue(
      "{AreaName}
      {lbl_value} (CI: {lbl_lcl} to {lbl_ucl})
      {lbl_num} out of a total of {lbl_denom} people
      "
    )
  ) |> 
  # plot as an interactive chart
  plotly::plot_ly(
    x = ~Value,
    y = ~AreaName,
    # name = ~AreaName,
    text = ~Text,
    hovertext = ~HoverText,
    hoverinfo = "text",
    type = "bar",
    marker = list(color = "#16a085")
  ) |> 
    plotly::layout(
      # set the title to the indicator name
      title = ind_data |> dplyr::pull(IndicatorShortName) |> unique(),
      # clear the y-axis title as its unecessary
      yaxis = list(title = ""),
      # set the x-axis range and title
      xaxis = list(range = c(0, 100), title = "Indicator performance (percent)"),
      # set the font
      font = list(family = "sans-serif", size = 14),
      # adjust the margin to avoid the title getting clipped
      margin = list(t = 80),
      # change the colour of the hovertext box to add some contrast
      hoverlabel = list(bgcolor = "#ecf0f1")
    ) |> 
      # hide the mode bar
    plotly::config(displayModeBar = FALSE)
```

Why this matters:

-   **Clear comparison:** Regions are ordered by performance, making inequality visible at a glance.

-   **Actionable insight:** Regions with lower rates can be targeted for intervention.

-   **Reusable workflow:** Swap out `indicator_id = 32` for AF, cholesterol or other conditions to replicate the analysis.